\bf Estimating the number of clusters using Cross Validation

### 2016-04-23 New simulation/technometrics.tex
Add confusion matrix

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{graphicx,psfrag,epsf}
\usepackage{enumerate}
\usepackage{natbib}
\usepackage{url} % not crucial - just used below for the URL 
%---------We added packages---------------------
%\usepackage{epsf}
\usepackage{hyperref}
%\usepackage{graphicx}
\usepackage{subfigure}
%\usepackage[margin=2.5cm]{geometry}
%\usepackage{amsmath}
\usepackage{footnote}
\usepackage[bottom]{footmisc}
\usepackage[singlelinecheck=false]{caption}
%\usepackage{caption}
\usepackage{amsthm}
\usepackage{enumitem}
%\usepackage{graphics}
\usepackage{float}
\usepackage{amssymb}
%\usepackage{natbib}
\usepackage{lscape}
\usepackage{array}
%\textheight=8.7in
\usepackage{setspace}
\usepackage{verbatim}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{longtable}

@@

It turns out that each cluster contains genes that tend to participate in the common processes. It is easy to see that cluster $5$ and $3$ are periodic from Figure \ref{fig:5_clusters}. Indeed, more than $70\%$ genes which function in mitotic cell cycle process are concentrated in these two clusters. Genes participated in the cell-cycle phase related processes such as DNA replication and DNA repair are highly concentrated in cluster $5$. Genes related to budding and cell polarity, as well as chromatin organization and transcription are commonly found in cluster $3$. These cyclic gene clusters are also reported in \cite{tavazoie1999systematic} and \cite{dortet2008model}. In fact, the mean expression profile plot of cluster $5$ resembles those of cluster $2$, $14$ of \cite{tavazoie1999systematic} and cluster $2$, $3$ of \cite{dortet2008model}, while cluster $3$ corresponding to cluster $7$ of \cite{tavazoie1999systematic}.

Cluster $1$ concentrates genes that somatize the cell stress, such as oxidative heat-induce proteins. This cluster corresponding to the cluster $5$ and $11$ in \cite{dortet2008model}.
Our analysis also found genes govern mitochondrial translation and mitochondrion organization are heavily concentrated in cluster $2$, while genes related to cytoplasmic translation are highlighted in cluster $4$. Genes encode ribosomes are also overly presented in cluster $4$. Those results are not reported in previous studies, meaning our analysis leads to new discovery of gene grouping.

In conclusion, functional grouping of gene occurred in the $5$ clusters found by our proposed method. In particular, the periodic cell cycle related genes are concentrated in cluster $5$ and cluster $3$, which broadly consistent with the findings in \cite{tavazoie1999systematic} and \cite{dortet2008model}. The number of clusters found by our method is much smaller than the number of clusters used in \cite{tavazoie1999systematic} and \cite{dortet2008model}. However, both studies found that most of their clusters have no characteristics, suggesting overestimate of the underlying diversity of biological expression classes in the data set \citep{tavazoie1999systematic}. Our result is more compact, while still able to identify the functional grouping of genes in clusters highlighted in previous studies.

@@

>>>>

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{graphicx,psfrag,epsf}
\usepackage{enumerate}
\usepackage{natbib}
\usepackage{url} % not crucial - just used below for the URL 
%---------We added packages---------------------
%\usepackage{epsf}
\usepackage{hyperref}
%\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{threeparttable}
%\usepackage[margin=2.5cm]{geometry}
%\usepackage{amsmath}
\usepackage{footnote}
\usepackage[bottom]{footmisc}
\usepackage[singlelinecheck=false]{caption}
%\usepackage{caption}
\usepackage{amsthm}
\usepackage{enumitem}
%\usepackage{graphics}
\usepackage{float}
\usepackage{amssymb}
%\usepackage{natbib}
\usepackage{lscape}
\usepackage{array}
%\textheight=8.7in
\usepackage{setspace}
\usepackage{verbatim}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{longtable}

@@

It turns out that each cluster contains genes that tend to participate in the common processes. It is easy to see that cluster $5$ and $3$ are periodic from Figure \ref{fig:5_clusters}. Indeed, more than $70\%$ genes which function in mitotic cell cycle process are concentrated in these two clusters. Genes participated in the cell-cycle phase related processes such as DNA replication and DNA repair are highly concentrated in cluster $5$. Genes related to budding and cell polarity, as well as chromatin organization and transcription are commonly found in cluster $3$. These cyclic gene clusters are also reported in \cite{tavazoie1999systematic} and \cite{dortet2008model}.

Cluster $1$ concentrates genes that somatize the cell stress, such as oxidative heat-induce proteins. This cluster corresponding to the cluster $5$ and $11$ in \cite{dortet2008model}.
Our analysis also found genes govern mitochondrial translation and mitochondrion organization are heavily concentrated in cluster $2$, while genes related to cytoplasmic translation are highlighted in cluster $4$. Genes encode ribosomes are also overly presented in cluster $4$.

Our results are well align with the results in \cite{tavazoie1999systematic}, which clustered the data with $k=30$. We prepared a confusion matrix (Table \ref{table:confusion}) to compare our results with those in \cite{tavazoie1999systematic}. In specific, \cite{tavazoie1999systematic} profiled three periodic clusters: cluster $2$, cluster $7$ and cluster $14$, which corresponding to our cluster $3$ and $5$. Indeed, we can easily see from Table \ref{table:confusion} that their cluster $2$ is entirely contained in our cluster $5$, their cluster $7$ is contained in our cluster $3$ and the periodic cluster $14$ is divided into our cluster $3$ and $5$. So their periodic clusters coincide with ours.

In addition, they also profiled several non-periodic clusters. Among those, their most notable functional grouping cluster $1$ is well contained inside our cluster $4$, their cluster $3$ is also in our cluster $3$ and their cluster $8$ is contained in our cluster $2$

In conclusion, functional grouping of gene occurred in the $5$ clusters found by our proposed method. In particular, the periodic cell cycle related genes are concentrated in cluster $5$ and cluster $3$, which consistent with the findings in \cite{tavazoie1999systematic} and \cite{dortet2008model}. The number of clusters found by our method is much smaller than the number of clusters used in \cite{tavazoie1999systematic} and \cite{dortet2008model}. However, both studies found that most of their clusters have no characteristics, suggesting overestimate of the underlying diversity of biological expression classes in the data set \citep{tavazoie1999systematic}. Our result is more compact, while still able to identify the functional grouping of genes in clusters highlighted in previous studies.

@@

\begin{table}[H]
\begin{center}
\captionsetup{justification=centering}
\caption{\label{table:confusion} Confusion matrix comparing $5$ clusters to $30$ clusters }
\begin{threeparttable}
\begin{tabular}{lccccccc}
\hline
   &  Cluster $1$  & Cluster $2$   &  Cluster $3$  &  Cluster $4$ &  Cluster $7$  & Cluster $8$  & Cluster $14$ \\ \hline 
 Cluster $1$  & $0$  &  $1$  & $0$  &  $0$  & $1$  & $3$  & $0$ \\
 Cluster $2$  & $0$  &  $0$  & $0$  & $102$ & $10$ & $145$& $1$ \\
 Cluster $3$  & $1$  &  $0$  & $91$ &  $2$  & $83$ & $0$  & $29$ \\
 Cluster $4$  & $161$&  $0$  & $11$ &  $66$ & $7$  & $0$  & $6$ \\
 Cluster $5$  & $2$  & $185$ & $2$  &  $0$  & $0$  & $0$  & $38$ \\ \hline      
\end{tabular}
 \begin{tablenotes}[para,flushleft]
\footnotesize {Only the profiled clusters in \cite{tavazoie1999systematic} are listed here}.
  \end{tablenotes}
\end{threeparttable}
\end{center}
\end{table}
